---
title: "Wrangle Google trend data"
author: "John Doe"
output: github_document
---

```{r setup, include=FALSE}
# create image folder ----
if (!file.exists("figs/")) {
    dir.create("figs/")
}
# create data folder ----
if (!file.exists("data/")) {
    dir.create("data/")
}
knitr::opts_chunk$set(
    echo = TRUE, # show all code
    eval = TRUE,
    message = FALSE, 
    comment = "#> ",
    warning = TRUE,
    tidy = FALSE, # cleaner code printing
    size = "small",
    fig.path = "figs/") # smaller code
knitr::opts_knit$set(
    width = 78)
base::options(tibble.print_max = 25,
              tibble.width = 78)
```


## Packages

The Google search trends are accessible via the [gtrendsR](https://github.com/PMassicotte/gtrendsR) package. This and other packages are in the `00-packages.R` script. We load these packages in the code chunk below. 

```{r 00-packages.R, eval=TRUE}
fs::dir_tree(".")
library(gtrendsR)
library(maps)
library(ggplot2)
library(lettercase)
library(viridis)
library(pals)
library(scico)
library(ggrepel)
library(tidyverse)
library(skimr)
```

## Import data

```{r import-data}
# import the 2019-06-27 data -------
# import the 2019-06-27 data using the script below
Dems2020Group1 <- readr::read_rds("data/raw/2019-06-28-Dems2020Group1.rds")
Dems2020Group2 <- readr::read_rds("data/raw/2019-06-28-Dems2020Group2.rds")

# import the 2019-06-28 data -------
# import the 2019-06-28 data using the script below
Dems2020Round2Group1 <- readr::read_rds("data/raw/2019-06-28-Dems2020Round2Group1.rds")
Dems2020Round2Group2 <- readr::read_rds("data/raw/2019-06-28-Dems2020Round2Group2.rds")

# import poll data --------------------------------------------------------
# fs::dir_tree("data")
RCPollData <- readr::read_csv("data/2019-06-28-RCPollData.csv")
```

### A little R programming

Your code will always be communicating to at least two audiences: your computer, and your future self. Be nice to both of them!

Things like the pipe `%>%` in R can help with clarity. It takes code written like this:

```r
outer_function(inner_function(Data_X), Data_Y)
```

And makes it look like this:

```r
Data_X %>% # do this 
   inner_function() %>% # then do this
   outer_function(Data_Y)
```

Convert to tibble. 

```{r Dems2020Group1, message=FALSE, warning=FALSE}
# convert Dems2020Group1 to tibble
Dems2020Group1IOT <- Dems2020Group1$interest_over_time %>% as_tibble()
# convert Dems2020Group2 to tibble
Dems2020Group2IOT <- Dems2020Group2$interest_over_time %>% as_tibble()
Dems2020Group1IOT %>% glimpse(78)
Dems2020Group2IOT %>% glimpse(78)
```

Change the `hits` to numeric (from string)

```{r numeric-hits, message=FALSE, warning=FALSE}
# create numeric hits 
Dems2020Group1IOT <- Dems2020Group1IOT %>% 
  dplyr::mutate(hits = as.numeric(hits)) 
Dems2020Group2IOT <- Dems2020Group2IOT %>%
  dplyr::mutate(hits = as.numeric(hits)) 
Dems2020Group1IOT %>% glimpse(78)
Dems2020Group2IOT %>% glimpse(78)
```

Bind these together. 

```{r Dems2020Debate01IOT}
# bind
Dems2020Debate01IOT <- bind_rows(Dems2020Group1IOT, 
          Dems2020Group2IOT,
          .id = "data") 
```

Check the min and max for the date. 

```{r min-max}
# get min
min(Dems2020Debate01IOT$date)
# get max
max(Dems2020Debate01IOT$date)
```

This is measuring interest from  `r min(Dems2020Debate01IOT$date)` until `r max(Dems2020Debate01IOT$date)` the next day.

## Wrangle

We will perform a few functions to get these data ready for visualizations.

### Men and Women

Create a `gender` variable for each candidate. 

```{r assign-gender}
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(gender = case_when(
    stringr::str_detect(keyword, "Elizabeth Warren") ~ "Women", 
    stringr::str_detect(keyword, "Amy Klobuchar") ~ "Women",
    stringr::str_detect(keyword, "Tulsi Gabbard") ~ "Women",
    TRUE ~ "Men"))
# get distinct
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% distinct()
# check new gender variable 
Dems2020Debate01IOT %>% 
  count(keyword, gender) %>%
  tidyr::spread(gender, n)
```


### Prior: Calculate the % of likely Democratic voters 

We will split the Men into three groups based on the [article updated on JUN. 26, 2019, AT 5:53 PM by fivethirtyeight](https://projects.fivethirtyeight.com/democratic-debate-poll/), "Who likely Democratic voters would vote for if the election were held today."

These groups were: `"> 1.0% of voters"`, `"0.5 - 0.9% of voters"`, `"0.2 - 0.4% of voters"`, and `"0.2 % of voters"`

See the categories below. 

```{r Dems2020Debate01IOT-prior_vperc}
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(prior_vperc = case_when(
    stringr::str_detect(keyword, "O’Rourke") ~ "> 1.0% of voters",
    stringr::str_detect(keyword, "Warren") ~ "> 1.0% of voters",
    stringr::str_detect(keyword, "Booker") ~ "> 1.0% of voters",
    
    stringr::str_detect(keyword, "Klobuchar") ~ "0.5 - 0.9% of voters",
    stringr::str_detect(keyword, "Castro") ~ "0.5 - 0.9% of voters",


    stringr::str_detect(keyword, "Gabbard") ~ "0.2 - 0.4% of voters",
    stringr::str_detect(keyword, "Ryan") ~ "0.2 - 0.4% of voters",
    stringr::str_detect(keyword, "Inslee") ~ "0.2 - 0.4% of voters",
    stringr::str_detect(keyword, "de Blasio") ~ "0.2 - 0.4% of voters",

    stringr::str_detect(keyword, "Delaney") ~ "0.2% of voters"))
Dems2020Debate01IOT %>% 
  count(prior_vperc, keyword) %>%
  tidyr::spread(prior_vperc, n)
```

Assign labels here. 

```{r prior_vperc_fct}
# assing labels
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(prior_vperc_fct = factor(x = prior_vperc))
# check levels 
dput(Dems2020Debate01IOT$prior_vperc_fct %>% levels())
```

```{r relevel}
Dems2020Debate01IOT <- Dems2020Debate01IOT %>% 
  dplyr::mutate(prior_vperc_fct = forcats::fct_relevel(.f = prior_vperc_fct,
                    "> 1.0% of voters",
                    "0.5 - 0.9% of voters",
                    "0.2 - 0.4% of voters",
                    "0.2% of voters"))

# check levels 
Dems2020Debate01IOT$prior_vperc_fct %>% levels()
```

```{r export}
# export Dems2020Debate01IOT data
readr::write_csv(as.data.frame(Dems2020Debate01IOT), paste0("data/",
                                        noquote(lubridate::today()),
                                        "-Dems2020Debate01IOT.csv"))
```


## Wrangle map data

What does the interest look like by location (i.e. `"region"`)? We can do a little wrangling and answer this question.

```{r WomanDems2020InterestByRegion}
# convert to tibble (another data structure in R)
Dems2020IBRGroup1 <- tibble::as_tibble(Dems2020Group1$interest_by_region)
Dems2020IBRGroup2 <- tibble::as_tibble(Dems2020Group2$interest_by_region)
# bind these together 
Dems2020IBR <- bind_rows(Dems2020IBRGroup1, 
                              Dems2020IBRGroup2, .id = "data")
Dems2020IBR %>% count(keyword)
```



Now we do a little wrangling on the `location` and `"state"`

```{r region}
# convert the region to lowercase
Dems2020InterestByRegion <- Dems2020IBR %>% 
  dplyr::mutate(region = stringr::str_to_lower(location))
# create a data set for the states in the US
statesMap = ggplot2::map_data("state")
# now merge the two data sources together
Dems2020InterestByRegion <- Dems2020InterestByRegion %>% 
  dplyr::inner_join(x = .,
                   y = statesMap, 
                   by = "region")
```


We've created a new data structure. What does it look like?

```{r Dems2020InterestByRegion}
# recheck the structure
Dems2020InterestByRegion %>%
  skimr::skim_to_wide() %>% 
  dplyr::filter(type %in% c("integer", "numeric")) %>% 
  dplyr::select(variable, 
                n, 
                mean, 
                sd, 
                median = p50, 
                hist)
```


It looks like the variable of interest (`hits`) is pretty lopsided--what can we do about it? After googling, we discover [this article](https://medium.com/@TheDataGyan/day-8-data-transformation-skewness-normalization-and-much-more-4c144d370e55) with this information, 

> "The logarithm, x to log base 10 of x, or x to log base e of x (ln x), or x to log base 2 of x, is a strong transformation and can be used to reduce right skewness."

### Export Map data

```{r export-Dems2020InterestByRegion}
# export Dems2020Debate01IOT data
readr::write_csv(as.data.frame(Dems2020InterestByRegion), paste0("data/",
                                        noquote(lubridate::today()),
                                        "-Dems2020InterestByRegion.csv"))
```

